---
title: "BAWL-R Stimuli Selection"
author: "Thomas Verliefde"
date: "2019-01-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
require(readxl)
require(stringr)
require(tidyverse)
require(tidyselect)
require(magrittr)
```

########################
# Exploring BAWL-R.xls #
########################

Critical Note:
For this code to work, you have to have the BAWL-R.xls file.
You can request access for research purposes from:
  http://www.ewi-psy.fu-berlin.de/einrichtungen/arbeitsbereiche/allgpsy/forschung/Download/index.html

```{r import, include=FALSE}

bawlR <- "BAWL-R.xls" %>% read_excel %>%
  mutate(
    WORD_CLASS = recode(
      WORD_CLASS,
      "N" = "NOUN",
      "V" = "VERB",
      "A" = "ADJ"
    )
  )

```

```{r explore letter range, include=FALSE}

bawlR %>% mutate(LETTERS = case_when(
    LETTERS <= 4 ~ "[,4]",
    LETTERS <= 7 ~ "[5,7]",
    LETTERS >= 8 ~ "[8,]"
  )) %>% count(WORD_CLASS,LETTERS)

```

```{r valence, options}

bawlR %>% group_by(WORD_CLASS) %>% summarize(min(EMO_MEAN),max(EMO_MEAN),sd(EMO_MEAN))

```





```{r previously used stims, options}

nounsAmb <- c("karriere","regulierung","arztbesuch","protest","alkohol","feuer","smartphone","stolz")

old <- bawlR %>% filter(
  WORD_LOWER %in% c(nounsAmb)
)
```

Note, 5 ambivalent words previously used by Berger are in the BAWLR, 3 are not.
But, both "arzt" and "besuch" are available (while "arztbesuch" is not)


```{r old, options}

old %>% select(WORD,`Ftot/1MIL`) %>% arrange(`Ftot/1MIL`)

old %>% select(WORD,EMO_MEAN,AROUSAL_MEAN,`Ftot/1MIL`,LETTERS) %>% t

bawlR %>% filter(WORD == "ARZT" | WORD == "BESUCH") %>%
  select(WORD,EMO_MEAN,AROUSAL_MEAN,`Ftot/1MIL`,LETTERS) %>% t

```

We will select 6 ambivalent words (the 5 available in the bawlR & "arztbesuch")
we will select 6 matching nounPos and 6 matching nounNeg (on letters and frequency).
-> Letters: 5,5,7,7,8,10


```{r selecting items, options}

adjPos <- bawlR %>% filter(
  WORD_CLASS == "ADJ",
  EMO_MEAN >= 2,
  `Ftot/1MIL` >= 10
) %T>% {
  count(.,LETTERS) %>% print
}

adjNeg <- bawlR %>% filter(
  WORD_CLASS == "ADJ",
  EMO_MEAN <= -2,
  `Ftot/1MIL` >= 10
) %T>% {
  count(.,LETTERS) %>% print
}

nounPos <- bawlR %>% filter(
  WORD_CLASS == "NOUN",
  EMO_MEAN >= 2,
  `Ftot/1MIL` >= 10,
  `Ftot/1MIL` <= 80,
  (LETTERS == 5 | LETTERS == 7 | LETTERS == 8 | LETTERS == 10)
) %T>% {
  count(.,LETTERS) %>% t %>% print
}

nounNeg <- bawlR %>% filter(
  WORD_CLASS == "NOUN",
  EMO_MEAN <= -2,
  `Ftot/1MIL` >= 10,
  `Ftot/1MIL` <= 80,
  (LETTERS == 5 | LETTERS == 7 | LETTERS == 8 | LETTERS == 10)
) %T>% {
  count(.,LETTERS) %>% t %>% print
}
```

#####################################################################
# Selecting/Creating Primes and Targets for Experiment and Practice #
#####################################################################

```{r creating random letter strings textfiles, options}

"Stimuli" %>% dir.create(showWarnings = FALSE)

# Strings of different sizes

set.seed(20190111); lapply(
  c(5,5,7,7,8,10)-1,
  function (x) {
      sample(LETTERS[-c(1,5,9,15,21,25)],1) %>% paste0(
        sample(letters[-c(1,5,9,15,21,25)],x, replace=FALSE) %>% paste(collapse="")
      )
  }
) %>% unlist %>% paste(collapse=" ") %>% writeLines("Stimuli/experimentPrime_Str.txt",sep="")

set.seed(20190111); lapply(
  c(5,7),
  function (x) {
      sample(LETTERS[-c(1,5,9,15,21,25)],1) %>% paste0(
        sample(letters[-c(1,5,9,15,21,25)],x, replace=FALSE) %>% paste(collapse="")
      )
  }
) %>% unlist %>% paste(collapse=" ") %>% writeLines("Stimuli/practicePrime_Str.txt",sep="")

```

```{r creating experiment primes and targets textfiles, options}
# Pos & Neg Nouns, in binsizes: 3-4, 5-6, 7-8, 9-10

"Stimuli" %>% dir.create(showWarnings = F)

primeAmb <- old %>% arrange(LETTERS) %>%
  add_row(WORD_LOWER = "arztbesuch", LETTERS = 10) %T>%
  {paste(.$WORD_LOWER,collapse=" ") %>% str_to_title %>%
      writeLines("Stimuli/experimentPrime_Amb.txt",sep="")}
  
primePos <- nounPos %>% filter(LETTERS == 5) %>%
  filter(
    WORD == "BLUME" |
      WORD == "HUMOR"
  ) %>% full_join(
    nounPos %>% filter(LETTERS == 7) %>%
      filter(
        WORD == "KOMFORT" |
          WORD == "TRIUMPH"
      ),by = vars_select(names(.),everything())
  ) %>% full_join(
    nounPos %>% filter(LETTERS == 8) %>%
      filter(
        WORD == "FREIZEIT"
      ),by = vars_select(names(.),everything())
  ) %>% full_join(
    nounPos %>% filter(LETTERS == 10) %>%
      filter(
        WORD == "LANDSCHAFT"
      ),by = vars_select(names(.),everything())
  ) %>% arrange(LETTERS) %T>% {paste(.$WORD_LOWER,collapse=" ") %>% str_to_title %>%
      writeLines("Stimuli/experimentPrime_Pos.txt",sep="")}

primeNeg <- nounNeg %>% filter(LETTERS == 5) %>%
  filter(
    WORD == "FEIND" |
      WORD == "RACHE"
  ) %>% full_join(
    nounNeg %>% filter(LETTERS == 7) %>%
      filter(
        WORD == "DROHUNG" |
          WORD == "DEFIZIT"
      ),by = vars_select(names(.),everything())
  ) %>% full_join(
    nounNeg %>% filter(LETTERS == 8) %>%
      filter(
        WORD == "ABSCHIED" 
      ),by = vars_select(names(.),everything())
  ) %>% full_join(
    nounNeg %>% filter(LETTERS == 10) %>%
      filter(
        WORD == "VERBRECHEN"
      ),by = vars_select(names(.),everything())
  ) %>% arrange(LETTERS) %T>% {paste(.$WORD_LOWER,collapse=" ") %>% str_to_title %>%
      writeLines("Stimuli/experimentPrime_Neg.txt",sep="")}

# Positive and Negative Targets, 8 each, only 4 each will be used

targetPos <- adjPos %>%
  filter(
    LETTERS >= 4,
    EMO_MEAN > mean(EMO_MEAN),
    WORD != "POSITIV"
  ) %T>% {paste(.$WORD,collapse=" ") %>%
        writeLines("Stimuli/experimentTarget_Pos.txt",sep="")}

targetNeg <- adjNeg %>%
  filter(
    LETTERS >= 4,
    WORD != "NEGATIV",
    WORD != "NUKLEAR",
    WORD != "SCHLECHT"
  ) %T>% {paste(.$WORD,collapse=" ") %>%
      writeLines("Stimuli/experimentTarget_Neg.txt",sep="")}


```

```{r creating practice prime and target textfiles, options}

"Stimuli" %>% dir.create(showWarnings = F)


# All identical to impamb2's practice words

paste("Kuss","Idee","Chance",collapse = " ") %>%
  writeLines("Stimuli/practicePrime_Pos.txt",sep="")

paste("Mord","Angst","Sklave",collapse = " ") %>%
  writeLines("Stimuli/practicePrime_Neg.txt",sep="")

paste("Geduldig","Zärtlich","Humorvoll","Fleissig",collapse = " ") %>%
  writeLines("Stimuli/practiceTarget_Pos.txt",sep="")

paste("Boshaft","Korrupt","Ungerecht","Gehässig",collapse = " ") %>%
  writeLines("Stimuli/practiceTarget_Neg.txt",sep="")

```

#####################################
# Analysing Chosen Primes & Targets #
#####################################
